Geometric variational problems

### 2020-11-12 NSF2020/Description/background.tex
edited Sections 1-3

\noindent
{\bf Collaborative Research: Mathematical modeling and simulations of
self-assembling amphiphilic particles in a solvent} \\
%{\bf {Collaborative Research: Mathematical modeling and coarse-grained simulations of self-assembly of amphiphilic Janus particles in a solvent}}\\
{\em Rolf Ryham (lead PI, Fordham University), Bryan Quaife (PI,
FSU), and Y.-N. Young (PI, NJIT)}
\section{Background}
\label{sec:background}
The goal of this collaborative proposal is to use mathetical modeling and
numerical simulations to investigate the dynamic self-assembly of
amphiphilic particles interacting with each other via hydrophobic forces
in a solvent. Amphiphilic particles (such as lipid molecules) possess
both hydrophobic and hydrophilic structures. In a viscous solvent they
self-assemble into meso-/macroscopic structures (such as micelles and
bilayers of lipids) to shield their hydrophobic parts from contact with
the solvent molecules (water).
%
%\subsection{Hydrophobic Forces}
%\label{sec:hydrophobicforce}
%
Such self-assembly of amphiphiles via hydrophobic force is ubiquitous in biology and biophysics \cite{Israelachvili1954},
%The hydrophobic force is a ubiquitous molecular interaction in biology \cite{Israelachvili1954}, 
and has been a major source of nonspecific interactions between
nanoparticles in  soft matter
\cite{Sanchez-IglesiasEtAl2012_ACSNano,AltantzisEtAl2013_PSC,XieYangLuEtAl2020_COCIS}. 
%The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles so to design smart materials of desirable properties by
%tuning the geometry and properties of the amphiphilic particles.

@@

\begin{wrapfigure}[10]{r}{0.22\textwidth}
\centerline{\includegraphics[width=0.22\textwidth]{figures/BG_fig1.pdf}}
\caption{ \footnotesize
A collection $P_1,$ $P_2,$ $P_3,$ \dots of rigid particles.
The exterior domain $\Omega$ represents the solvent.}
  \label{fig:domain}
\end{wrapfigure}
Recently PIs RR and YNY developed a mathematical model called the
hydrophobic attraction potential (HAP) \cite{Fu2018_SIAM} that is based
on the physical origin of hydrophobicity.  This model addresses the
major shortcomings of molecular dynamics (MD) and continuum approaches.
Based on preliminary results (\S\ref{sec:preliminary_work}) the PIs
propose to extend this HAP model to offer an alternative modeling
methodology that leads to new mathematical ideas, and is both physically
accurate and computationally practical.
%
%
%The word hydrophobic (water fearing) derives from the low solubility of oil (hydrocarbon solute) in water and vice versa. 
%It causes hydrophobic moieties to aggregate and cluster,
%is responsible for the adhesion between hydrophobic surfaces \cite{Ducker2016}, large contact angles on a 
%dewetting surface \cite{Arenas2019,Sandre1999}, accumulation of particles along interfaces \cite{Lee2013,Lee2014}, 
%formation of micelles and bilayers \cite{Israelachvili80}, and protein folding and membrane insertion \cite{Kabelka2018}.
%
%Lipids are amphiphilic molecules whose  structure possesses
%both hydrophobic and hydrophilic parts. 
%The amphiphilic property is what allows lipids to form the membranes and 
%compartments of living cells \cite{Israelachvili80}. 
%More specifically, a lipid consists of 
%an elongated hydrocarbon tail that is hydrophobic, attached to a polar head that is hydrophilic.
%To shield the hydrophobic tails from water, lipids self-assemble into micelles and bilayers. 
%A micelle is a spherical arrangement of lipids with tails terminating at the micelle center. 
%A bilayer consists of two layers of lipids called monolayers, where the lipid tails point 
%from the monolayer surface into the bilayer core. 
%
%The mathematical modeling of a biological membrane is a challenging problem in applied mathematics. 
%Bilayers are elastic and resist deformations like bending, twisting, and stretching.
%Their elastic deformations are well described by the theory of liquid crystals \cite{ANDRIENKO2018520}.
%Lipid bilayer membranes can also be fluidic, and the lateral translation of lipids (or any membrane bound
%proteins) couples nonlocaly to the motion of the aqueous environment \cite{MerkelSackmannEvans1989,StoneAjdari1998_JFM,OppenheimerDiamant2009_BJ,OppenheimerDiamant2011_PRL}. 
%%Finally, the membranes of cellular 
%%compartments are constantly merging and pinching as part of intracellular trafficking. 
%%Therefore monolayer surfaces undergo discontinuous deformations. 
%
%There are two prevailing approaches in membrane modeling, and each has its advantages
%and disadvantages. Molecular dynamics (MD) is to date the only tool capable of resolving granular biological details
%The computational cost of MD, however, grows with the sixth power of the sample diameter and 
%so simulations are severely limited to 
%small system sizes and short time scales \cite{DiCarlo2019}.
%The other approach, continuum mechanics, assumes smooth surfaces and can therefore model 
%realistic systems over physical times. 
%But continuum description of a biological membrane ignores the granularity of lipid molecules, and thus requires some assumptions when a
%membrane ruptures or when two membranes fuse \cite{ChKo08}. 
%%
%%mechanics presumes, rather than predicts, the nucleation of discontinuities. 
%%
%
The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles. These results will facilitate optimal design of smart materials by
tuning the geometry and properties of the amphiphilic particles.

@@

To define HAP, consider a collection of particles solvent suspended in a solvent. 
The region $\Omega \subset \mathbb{R}^3$  models 
the solvent phase (Figure \ref{fig:domain}).
%The HAP must be an energy of the shape of the region because hydrophobic
%attraction is non-additive.
The boundary of the region is the
water-particle interface.
Some parts of this interface are hydrophobic
while others are hydrophilic.
Hydrophobic interfaces disturb the hydrogen bond structure of water~\cite{Luzar1987,
Jonsson2006, Varilly2011}. This disturbance comes with an energetic penalty
that is proportional to interfacial area.
There is an additional decay length,
due to rapid fluctuation in the hydrogen bond
network, describing how far the restructuring extends into bulk water.

@@

To get the effect of attraction, the water activity is not arbitrary, but rather
is the unique solution of the screened Laplace equation boundary value problem 
\begin{equation}
  \label{SL}
  -\rho^2 \Delta u + u = 0, \mbox{ } \mathbf{x} \in \Omega, \qquad
  u = f,  \mbox{ } \mathbf{x} \in \partial \Omega, \qquad 
  u(\mathbf{x}) \to 0 \mbox{ as } |\mathbf{x}| \to \infty.
\end{equation}
The boundary values $f$ define the degree of hydrophobicity of the
water-particle interface: $f=1$, describes a hydrophobic interface
and $f=0$ describes a hydrophilic interface. Thus $f$ encodes information
about the particles and the parameters $\rho$ and $\gamma$ encode information about the quality of the solvent 
~\cite{Israelachvili1954, Discher2002}. In practice, we use
\eqref{SL} to integrate by parts to obtain
\begin{equation}
\label{eq:HAP_easy}
\Phi = -\gamma \int_{\partial \Omega} \rho u \nabla u \cdot \boldsymbol{\nu} \,\dif S,
\end{equation}
thereby avoiding calculating a volumetric integral.

The equations~\eqref{HAP} and~\eqref{SL} possess number of the mathematical properties
that capture the phenomenological behavior of hydrophobic attraction.
Specifically, solutions of~\eqref{SL} yield an attractive interaction between
hydrophobic bodies separated by water ~\cite{Eriksson1989}.
Using boundary layer analysis, the HAP converges to a surface energy in the zero-decay length
limit~\cite{Lee2018, Lin2015, Shibata2004}.
Finally, we have demonstrated that the forces derived from HAP theory are non-additive~\cite{Meyer2006, Fu2018_SIAM}. 
%The hydrophobic force has been implicated in the directed folding of proteins, 
%adhesion between biological membranes, but it is still unanswered as to whether
%the hydrophobic force of the form (\ref{HAP}--\ref{SL}) exists between small molecules.

@@

To define the particle dynamics,
let $P_1$, $P_2$, \ldots, $P_N$,
be a finite collection of disjoint, rigid and closed particles
with Lipschitz boundary.
Assume that the label $f$ is an element of the Sobolev space $H^1(\Omega)$.
The functional~\eqref{HAP} has a minimizer among all functions $u$ equaling $f$ on the boundary in the sense of trace.
Conversely, from maximum principles and energy estimates, the solution of~\eqref{SL} is unique and minimizes~\eqref{HAP}.
Taking the first variation of~\eqref{HAP}, i.e. the derivative with respect to the domain
~\cite{Bandle2015, Schiffer1954, Grinfeld2010},
yields a symmetric, rank-two tensor called the hydrophobic stress:

@@

  (\boldsymbol{\sigma}_{\text{hydro}} \cdot \boldsymbol{\nu}) \,\dif S

@@

relative to the center of mass $\mathbf{a}_i$.
This system is force- and torque-free  \S\ref{subsec:specific_aim_2}.
To avoid particle collisions, we define an excluded volume potential $\Phi_{\text{repul}}$
that diverges whenever tubular neighbordhoods of adjacent particles overlap,
along with the excluded volume repulsions. The total potential, force and torque are then

@@

The time integration of
particle configurations goes as follows: \textbf{(i)} solve the
BVP~\eqref{SL} for the screened Laplace equation, \textbf{(ii)}
determine the rigid body forces and torques~\eqref{forceandtorque},
\textbf{(iii)} solve the Stokes mobility problem~\eqref{eq:stokes} for
rigid body motion, \textbf{(iv)} update the particle configuration.
Numerical challenges and how these challenges are overcome are described
in \S\ref{subsec:specific_aim_2}.

@@

The HAP formulation is, to our knowledge, the first demonstration of bilayer self-assembly
by a continuum-based interaction model~\cite{Noguchi2001, Farago2003, Brannigan2006,
Brooks2009, Wang2013}.
Our simulations use Janus particles to model lipid amphiphiles which are
popular in material science and physics for creating functional
materials~\cite{Lee2014, Lee2013}. Janus particles are typically
spherical with a biphasic material label on either hemisphere, endowing
the particle with a directional order.
We model elongated lipid by elliptical particles with the hydrophobic label defined  along the ellipse's axis.
Under the hydrophobic force, and with an excluded volume,
the Janus particles spontaneously merge and realign to form
bilayers. This occurs only as a result of energy minimization and does
not require artificial inputs.

@@

As a proof of concept, our work has already tested for elastic energies
for bending, stretching, and tilt of the bilayer assembly. 
The elastic coefficients derived from the HAP simulations show strikingly positive
agreement with experimentally determined values~\cite{Fu2018_SIAM}.
Encouraged by these results and the hydrodynamic simulations of the
bilayer assembly, the PIs propose to extend the HAP model
to make direct comparison with experimental results  \S\ref{subsec:specific_aim_1},
to establish that the HAP model has the capability
and adaptivity to model flow phenomena across length scales and time scales \S\ref{subsec:specific_aim_2},
using efficient, high-order numerical methods \S\ref{sec:proposed-work}.

\begin{wrapfigure}[14]{l}{0.35\textwidth}
\centerline{\includegraphics[width=0.35\textwidth]{figures/PW_fig2.pdf}}
  \vspace{-8pt}
  \caption{\label{fig:demixing} \footnotesize An initial assembly of
  small and large particles spontaneously segregates into two smaller
  bodies.}
\end{wrapfigure}Over the past decades, researchers have used a number of mathematical
tools to simulate vesicles in a shear flow, including lattice
Boltzmann~\cite{KaouiHartingMisbah2011_PRE}, coarse-grained Brownian
dynamics~\cite{NoguchiTakasu2002_BJ}, phase
field~\cite{DuLiuWang2004_JCP,BibenKassnerMisbah2005_PRE}, level
set~\cite{DoyeuxGuyotChabannesEtAl2013_JCAM}, boundary
integral~\cite{Shravan09,Rahimian15} and immersed boundary
approaches~\cite{KimLai2010_JCP,KimLai2012_PRE,HuLaiSeolEtAl2016_JCP}.
Most of these approaches assume a mathematical surface, whether
implicitly or explicitly, and define an elastic bending energy of the
surface. Researchers have developed a number of numerical methods for
calculating energy minimizing steady equilibrium shapes of lipid bilayer
membranes, vesicles and red blood cells. These approaches range from the
finite element method~\cite{Bartels,Peng13,RyKlYaCo16,Sinha15}, phase
field method~\cite{Du05,QiangDu08,Lowengrub13} to immersed boundary
method~\cite{Hu,Hu13, KimLai2010_JCP}. PI RR and collaborators led in
part the development of phase field functionals of membrane elastic
energy and approaches to coupling membrane elasticity to
fluids~\cite{0951-7715-18-3-016,Du05,DuEuler,QiangDu09}.

@@

Our HAP approach differs from these prior methods in a number of respects.
First, we do not assume a surface. Rather, we more fundamentally
consider a collection of amphiphilic particles.  The collection of
amphiphiles minimize hydrophobic interactions by sequestering
hydrophobic tails in the form of a bilayer, and the particles' excess
free energy gives rise to an elastic bilayer energy. The second
difference lies in the fluid-interface coupling. Here, the associated
mobility problem~\eqref{eq:stokes} is more complicated than dealing with
a stress boundary condition or diffusive surface force, because the
fluid velocity must be that of a rigid body motion at the particle
surfaces. The HAP model also addresses the existence of multiple phases.
We can vary lipid length, spontaneous curvature, and bending rigidity by
introducing different particle shapes and hydrophobic boundary
conditions (Figure~\ref{fig:demixing}). In contrast, continuum theory
deals with multiple phases through additional surface densities that
must then satisfy specialized transport equations~\cite{Lowengrub07,
MikuckiZhou17}.

The greatest strength of using HAP to model lipid bilayer membrane is
the ability to form discontinuities (interfacial singularity) from first
physical principles without any artificial manipulation to rearrangement
the interface. For example, using the HAP model to simulate a vesicle
under a shear flow that is strong enough to cause the vesicle membrane
to rupture, we expect the HAP model to capture the reorganization of
lipid molecules on the scales of membrane thickness ($\sim 5$ nm), which
is a nearly impossible task using phase-field or immerse boundary method
without extreme refinement around the membrane and artificial treatment
of reconnection during the topological change of an interface.

>>>>

\noindent
{\bf Collaborative Research: Mathematical modeling and simulations of
self-assembling amphiphilic particles in a solvent} \\
%{\bf {Collaborative Research: Mathematical modeling and coarse-grained simulations of self-assembly of amphiphilic Janus particles in a solvent}}\\
{\em Rolf Ryham (lead PI, Fordham University), Bryan Quaife (PI,
FSU), and Y.-N. Young (PI, NJIT)}
\section{Background}
\label{sec:background}
The goal of this collaborative proposal is to use mathematical modeling
and numerical simulations to investigate the dynamic self-assembly of
amphiphilic particles interacting with each other via hydrophobic forces
in a solvent. Amphiphilic particles (such as lipid molecules) possess
both hydrophobic and hydrophilic structures. In a viscous solvent they
self-assemble into meso-/macroscopic structures (such as micelles and
bilayers of lipids) to shield their hydrophobic parts from contact with
the solvent molecules (water).
%
%\subsection{Hydrophobic Forces}
%\label{sec:hydrophobicforce}
%
Such self-assembly of amphiphiles via hydrophobic forces is ubiquitous in biology and biophysics \cite{Israelachvili1954},
%The hydrophobic force is a ubiquitous molecular interaction in biology \cite{Israelachvili1954}, 
and has been a major source of nonspecific interactions between
nanoparticles in  soft matter
\cite{Sanchez-IglesiasEtAl2012_ACSNano,AltantzisEtAl2013_PSC,XieYangLuEtAl2020_COCIS}. 
%The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles so to design smart materials of desirable properties by
%tuning the geometry and properties of the amphiphilic particles.

@@

\begin{wrapfigure}[10]{r}{0.22\textwidth}
\centerline{\includegraphics[width=0.22\textwidth]{figures/BG_fig1.pdf}}
\caption{ \footnotesize
A collection $P_1,$ $P_2,$ $P_3, \ldots$ of rigid particles.  The
  exterior domain $\Omega$ represents the solvent.}
  \label{fig:domain}
\end{wrapfigure}
Recently PIs RR and YNY developed a mathematical model called the
hydrophobic attraction potential (HAP) \cite{Fu2018_SIAM} that is based
on the physical origin of hydrophobicity.  This model addresses the
major shortcomings of molecular dynamics (MD) and continuum approaches.
Based on preliminary results (\S\ref{sec:preliminary_work}) the PIs
propose to extend this HAP model to offer an alternative modeling
methodology that leads to new mathematical ideas, and is both physically
accurate and computationally practical.
%
%
%The word hydrophobic (water fearing) derives from the low solubility of oil (hydrocarbon solute) in water and vice versa. 
%It causes hydrophobic moieties to aggregate and cluster,
%is responsible for the adhesion between hydrophobic surfaces \cite{Ducker2016}, large contact angles on a 
%dewetting surface \cite{Arenas2019,Sandre1999}, accumulation of particles along interfaces \cite{Lee2013,Lee2014}, 
%formation of micelles and bilayers \cite{Israelachvili80}, and protein folding and membrane insertion \cite{Kabelka2018}.
%
%Lipids are amphiphilic molecules whose  structure possesses
%both hydrophobic and hydrophilic parts. 
%The amphiphilic property is what allows lipids to form the membranes and 
%compartments of living cells \cite{Israelachvili80}. 
%More specifically, a lipid consists of 
%an elongated hydrocarbon tail that is hydrophobic, attached to a polar head that is hydrophilic.
%To shield the hydrophobic tails from water, lipids self-assemble into micelles and bilayers. 
%A micelle is a spherical arrangement of lipids with tails terminating at the micelle center. 
%A bilayer consists of two layers of lipids called monolayers, where the lipid tails point 
%from the monolayer surface into the bilayer core. 
%
%The mathematical modeling of a biological membrane is a challenging problem in applied mathematics. 
%Bilayers are elastic and resist deformations like bending, twisting, and stretching.
%Their elastic deformations are well described by the theory of liquid crystals \cite{ANDRIENKO2018520}.
%Lipid bilayer membranes can also be fluidic, and the lateral translation of lipids (or any membrane bound
%proteins) couples nonlocaly to the motion of the aqueous environment \cite{MerkelSackmannEvans1989,StoneAjdari1998_JFM,OppenheimerDiamant2009_BJ,OppenheimerDiamant2011_PRL}. 
%%Finally, the membranes of cellular 
%%compartments are constantly merging and pinching as part of intracellular trafficking. 
%%Therefore monolayer surfaces undergo discontinuous deformations. 
%
%There are two prevailing approaches in membrane modeling, and each has its advantages
%and disadvantages. Molecular dynamics (MD) is to date the only tool capable of resolving granular biological details
%The computational cost of MD, however, grows with the sixth power of the sample diameter and 
%so simulations are severely limited to 
%small system sizes and short time scales \cite{DiCarlo2019}.
%The other approach, continuum mechanics, assumes smooth surfaces and can therefore model 
%realistic systems over physical times. 
%But continuum description of a biological membrane ignores the granularity of lipid molecules, and thus requires some assumptions when a
%membrane ruptures or when two membranes fuse \cite{ChKo08}. 
%%
%%mechanics presumes, rather than predicts, the nucleation of discontinuities. 
%%
%
The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles. These results will facilitate optimal design of smart materials by
tuning the geometry and properties of the amphiphilic particles.

@@

To define HAP, consider a collection of particles suspended in a
solvent. The region $\Omega \subset \mathbb{R}^3$ models the solvent
phase (Figure \ref{fig:domain}).
%The HAP must be an energy of the shape of the region because hydrophobic
%attraction is non-additive.
The boundary of the region is the water-particle interface with outward
pointing normal $\boldsymbol{\nu}$. Some parts of this interface are
hydrophobic while others are hydrophilic. Hydrophobic interfaces disturb
the hydrogen bond structure of water~\cite{Luzar1987, Jonsson2006,
Varilly2011}. This disturbance comes with an energetic penalty that is
proportional to interfacial area. There is an additional decay length,
due to rapid fluctuation in the hydrogen bond network, describing how
far the restructuring extends into bulk water.

@@

To establish the effect of attraction, the water activity is not
arbitrary, but rather is the unique solution of the screened Laplace
equation boundary value problem 
\begin{equation}
  \label{SL}
  -\rho^2 \Delta u + u = 0, \mbox{ } \mathbf{x} \in \Omega, \qquad
  u = f,  \mbox{ } \mathbf{x} \in \partial \Omega, \qquad 
  u(\mathbf{x}) \to 0 \mbox{ as } |\mathbf{x}| \to \infty.
\end{equation}
The boundary values $f$ define the degree of hydrophobicity of the
water-particle interface: $f=1$, describes a hydrophobic interface and
$f=0$ describes a hydrophilic interface. Thus $f$ encodes information
about the particles and the parameters $\rho$ and $\gamma$ encode
information about the quality of the solvent~\cite{Israelachvili1954,
Discher2002}. In practice, we integrate~\eqref{HAP} by parts and
use~\eqref{SL} to obtain
\begin{equation}
\label{eq:HAP_easy}
\Phi = -\gamma \int_{\partial \Omega} \rho u \nabla u \cdot \boldsymbol{\nu} \,\dif S,
\end{equation}
thereby avoiding calculating a volumetric integral.

The equations~\eqref{HAP} and~\eqref{SL} possess a number of the mathematical properties
that capture the phenomenological behavior of hydrophobic attraction.
Specifically, solutions of~\eqref{SL} yield an attractive interaction between
hydrophobic bodies separated by water~\cite{Eriksson1989}.
Using boundary layer analysis, the HAP converges to a surface energy in the zero-decay length
limit~\cite{Lee2018, Lin2015, Shibata2004}.
Finally, we have demonstrated that the forces derived from HAP theory are non-additive~\cite{Meyer2006, Fu2018_SIAM}. 
%The hydrophobic force has been implicated in the directed folding of proteins, 
%adhesion between biological membranes, but it is still unanswered as to whether
%the hydrophobic force of the form (\ref{HAP}--\ref{SL}) exists between small molecules.

@@

To define the particle dynamics, let $P_1$, $P_2$, \ldots, $P_N$, be a
finite collection of disjoint, rigid, and closed particles with
Lipschitz boundary. Assume that the label $f$ is an element of the
Sobolev space $H^1(\Omega)$. The functional~\eqref{HAP} has a minimizer
among all functions $u$ equaling $f$ on the boundary in the sense of
trace.  Conversely, from maximum principles and energy estimates, the
solution of~\eqref{SL} is unique and minimizes~\eqref{HAP}. Taking the
first variation of~\eqref{HAP}, i.e.~the derivative with respect to the
domain~\cite{Bandle2015, Schiffer1954, Grinfeld2010}, yields a
symmetric, rank-two tensor called the hydrophobic stress:

@@

  (\boldsymbol{\sigma}_{\text{hydro}} \cdot \boldsymbol{\nu}) \,\dif S,

@@

relative to the center of mass $\mathbf{a}_i$.  This system is force-
and torque-free (\S\ref{subsec:specific_aim_2}).  To avoid particle
collisions, we define an excluded volume potential $\Phi_{\text{repul}}$
that diverges whenever tubular neighborhoods of adjacent particles
overlap, along with the excluded volume repulsions. The total potential,
force and torque are then

@@

The time integration of
particle configurations goes as follows: \textbf{(i)} solve the
BVP~\eqref{SL} for the screened Laplace equation, \textbf{(ii)}
determine the rigid body forces and torques~\eqref{forceandtorque},
\textbf{(iii)} solve the Stokes mobility problem~\eqref{eq:stokes} for
rigid body motion, \textbf{(iv)} update the particle configuration.
Numerical challenges and how these challenges are overcome are described
in \S\ref{subsec:specific_aim_2} and \S\ref{subsec:specific_aim_3}.

@@

The HAP formulation is, to our knowledge, the first demonstration of
bilayer self-assembly by a continuum-based interaction
model~\cite{Noguchi2001, Farago2003, Brannigan2006, Brooks2009,
Wang2013}. Our simulations use Janus particles to model lipid
amphiphiles which are popular in material science and physics for
creating functional materials~\cite{Lee2014, Lee2013}. Janus particles
are typically spherical with a biphasic material label on either
hemisphere, endowing the particle with a directional order. We model an
elongated lipid by elliptical particles with the hydrophobic label
defined along the ellipse's axis. Under the hydrophobic force, and with
an excluded volume, the Janus particles spontaneously merge and realign
to form bilayers. This occurs only as a result of energy minimization
and does not require artificial inputs.

@@

As a proof of concept, our work has already tested for elastic energies
for bending, stretching, and tilt of the bilayer assembly.  The elastic
coefficients derived from the HAP simulations show strikingly positive
agreement with experimentally determined values~\cite{Fu2018_SIAM}.
Encouraged by these results and the hydrodynamic simulations of the
bilayer assembly, the PIs propose to extend the HAP model to make direct
comparison with experimental results  \S\ref{subsec:specific_aim_1}, to
establish that the HAP model has the capability and adaptivity to model
flow phenomena across length scales and time scales
\S\ref{subsec:specific_aim_2}, using efficient, high-order numerical
methods \S\ref{subsec:specific_aim_3}.

\begin{wrapfigure}[14]{l}{0.35\textwidth}
\centerline{\includegraphics[width=0.35\textwidth]{figures/PW_fig2.pdf}}
  \vspace{-8pt}
  \caption{\label{fig:demixing} \footnotesize An initial assembly of
  small and large particles spontaneously segregates into two smaller
  bodies.}
\end{wrapfigure}
Over the past decades, researchers have used a number of mathematical
tools to simulate vesicles in a shear flow, including lattice
Boltzmann~\cite{KaouiHartingMisbah2011_PRE}, coarse-grained Brownian
dynamics~\cite{NoguchiTakasu2002_BJ}, phase
field~\cite{DuLiuWang2004_JCP,BibenKassnerMisbah2005_PRE}, level
set~\cite{DoyeuxGuyotChabannesEtAl2013_JCAM}, boundary
integral~\cite{Shravan09,Rahimian15}, and immersed boundary
approaches~\cite{KimLai2010_JCP,KimLai2012_PRE,HuLaiSeolEtAl2016_JCP}.
Most of these approaches assume a mathematical surface, whether
implicitly or explicitly, and define an elastic bending energy of the
surface. Researchers have developed a number of numerical methods for
calculating energy minimizing steady equilibrium shapes of lipid bilayer
membranes, vesicles, and red blood cells. These approaches range from
the finite element method~\cite{Bartels,Peng13,RyKlYaCo16,Sinha15},
phase field method~\cite{Du05,QiangDu08,Lowengrub13}, and immersed
boundary method~\cite{Hu,Hu13, KimLai2010_JCP}. PI RR and collaborators
led in part the development of phase field functionals of membrane
elastic energy and approaches to coupling membrane elasticity to
fluids~\cite{0951-7715-18-3-016,Du05,DuEuler,QiangDu09}.

@@

Our HAP approach differs from these prior methods in a number of respects.
First, we do not assume a surface. Rather, we more fundamentally
consider a collection of amphiphilic particles.  The collection of
amphiphiles minimize hydrophobic interactions by sequestering
hydrophobic tails in the form of a bilayer, and the particles' excess
free energy gives rise to an elastic bilayer energy. The second
difference lies in the fluid-interface coupling. Here, the associated
mobility problem~\eqref{eq:stokes} is more complicated than dealing with
a stress boundary condition or diffusive surface force, because the
fluid velocity must be that of a rigid body motion at the particle
surfaces. The HAP model also addresses the existence of multiple phases.
We can vary lipid length, spontaneous curvature, and bending rigidity by
introducing different particle shapes and hydrophobic boundary
conditions (Figure~\ref{fig:demixing}). In contrast, continuum theory
deals with multiple phases through additional surface densities that
must satisfy specialized transport equations~\cite{Lowengrub07,
MikuckiZhou17}.

\begin{wrapfigure}[6]{l}{0.35\textwidth}
\centerline{\includegraphics[width=0.35\textwidth]{figures/Lysis.pdf}}
  \vspace{-8pt}
  \caption{\label{fig:lysis} \footnotesize Two hydrophobic particles
  enter and then lyse the circular bilayer.}
\end{wrapfigure}
The greatest strength of using HAP to model lipid bilayer membrane is
the ability to form discontinuities (interfacial singularity) from first
physical principles without any artificial manipulation to rearrange the
interface (Figures~\ref{fig:demixing} and~\ref{fig:lysis}).  For
example, using the HAP model to simulate a vesicle under a shear flow
that is strong enough to cause the vesicle membrane to rupture (see
\S\ref{sec:preliminary_work}), we expect the HAP model to capture the
reorganization of lipid molecules on the scales of membrane thickness
($\sim 5$ nm), which is a nearly impossible task using phase-field or
immerse boundary method without extreme refinement around the membrane
and artificial treatment of reconnection during the topological change
of an interface \cite{doi:10.1063/5.0009734, LiAn-Chang16,
doi:10.1098/rspa.2012.0505, doi:10.1137/130941432, Feetzl18,
doi:10.1137/16M1108406}.

### 2020-11-12 NSF2020/Description/broaderimpacts.tex
edited Sections 1-3

The proposed mathematical analysis, modeling, and numerical algorithms
will transform our understanding of the collective dynamics of
amphiphilic particles such as (1) their self-assembly into micelles and
bilayers, (2) the material properties of such self-assembly, and (3) the
interaction between these building blocks. In addition to biophysical
applications, amphiphilic Janus particles have recently popularity for
fabrication of smart materials. The proposed work will have a
transformative impact on precision design for specific mechanical
properties of materials made of amphiphilic nanoparticles. An important
component of this proposal is the interdisciplinary education and
training of both undergraduate students, graduate students, and
postdoctoral fellows. The combination of mathematical modeling,
analysis, and scientific computing in this project provides a compelling
example of the importance of mathematics in biophysics and engineering
applications. The concepts and methods described here go beyond the
context of amphiphilic particles and lipid molecules. They extend to
other problems featuring microscopic phase separation that leads to
formation of mesoscopic domains. This situation arises, for example, in
biological development in systems biology. The methods developed here
have the potential to impact those and other related areas in
biomedicine and biotechnology.

\subsection{Educational Impacts}
\label{subsec:Educational_plans}
The proposed research will have immediate impact on undergraduate and
graduate education. At the undergraduate level, the project will involve
and support undergraduate researcher from Fordham University. The
supported summer researchers will gain valuable first-hand
interdisciplinary experience in mathematical fluid modeling and
computation. We will incorporate research into teaching, especially in
differential equation and programming courses. PI YNY will work on
simple modeling of self-assembly problems with both undergraduate and
high school students from Newark Science Park High School. PI YNY has a
track record of working with local high school students for their
research projects before they apply for colleges. PI BQ will work with
undergraduate and high school students through the Undergraduate
Research Opportunity Program (UROP) and Young Scholar's Program (YSP) as
he has done in the past.

@@

\section{Intellectual Merit}
A central theme of the project is the mathematical development and
analysis of the physical model for interaction between many amphiphilic
particles. The model formulates the interaction potential through a
screened Laplace equation boundary value problem possessing the physical
properties like non-additivity and decay of realistic hydrophobic
attraction. Colloidal systems collectively self-assemble into bilayer
morphologies, and we analyze the elastic properties of these amphiphilic
particle ensembles. This allows us to interpret the Helfrich free energy
in terms of hydrophobic interactions and specific molecular
characteristics. We extend the capabilities of the boundary integral
equation and time stepping methods to stably perform large particle
number simulations in three dimensions. Results from these computations
will be used to compare collective amphiphilic against experiment, and
study the optimal design of three-dimensional functional materials.

>>>>

The proposed mathematical analysis, modeling, and numerical algorithms
will transform our understanding of the collective dynamics of
amphiphilic particles such as (1) their self-assembly into micelles and
bilayers, (2) the material properties of such self-assembly, and (3) the
interaction between these building blocks. In addition to biophysical
applications, amphiphilic Janus particles have recent popularity in the
fabrication of smart materials. The proposed work will have a
transformative impact on precision design for specific mechanical
properties of materials made of amphiphilic nanoparticles. An important
component of this proposal is the interdisciplinary education and
training of both undergraduate students, graduate students, and
postdoctoral fellows. The combination of mathematical modeling,
analysis, and scientific computing in this project provides a compelling
example of the importance of mathematics in biophysics and engineering
applications. The concepts and methods described here go beyond the
context of amphiphilic particles and lipid molecules. They extend to
other problems featuring microscopic phase separation that leads to
formation of mesoscopic domains. This situation arises, for example, in
biological development in systems biology. The methods developed here
have the potential to impact those and other related areas in
biomedicine and biotechnology.

\subsection{Educational Impacts}
\label{subsec:Educational_plans}
The proposed research will have an immediate impact on undergraduate and
graduate education. At the undergraduate level, the project will include
and support undergraduate researchers from Fordham University. The
supported summer researchers will gain valuable first-hand
interdisciplinary experience in mathematical fluid modeling and
computation. We will incorporate research into teaching, especially in
differential equation and programming courses. PI YNY will work on
simple modeling of self-assembly problems with both undergraduate and
high school students from Newark Science Park High School. PI YNY has a
track record of working with local high school students for their
research projects before they apply for colleges. PI BQ will work with
undergraduate and high school students through the Undergraduate
Research Opportunity Program and Young Scholar's Program as he has done
in the past.

@@

\section{Intellectual Merit}
A central theme of the project is the mathematical development and
analysis of the physical model for interaction between many amphiphilic
particles. The model formulates the interaction potential through a
screened Laplace equation boundary value problem possessing the physical
properties like non-additivity and decay of realistic hydrophobic
attraction. Colloidal systems collectively self-assemble into bilayer
morphologies, and we analyze the elastic properties of these amphiphilic
particle ensembles. This allows us to interpret the Helfrich free energy
in terms of hydrophobic interactions and specific molecular
characteristics. We extend the capabilities of the boundary integral
equation and time stepping methods to stably perform large particle
number simulations in three dimensions. Results from these computations
will be used to compare collective amphiphilic against experiments, and
study the optimal design of three-dimensional functional materials.
